# AKSTA Quiz Notes 3

🤓 Let's try and have some fun with Data Science 🤓

```{r}
demo("graphics")
```

Okay that was actually fun 🙂

## What are the three groups that R plotting functions can be divided into? ✅

* High-level --> create new plots (sometimes as objects)
* Low-level --> add information to an existing plot
* Interactive graphics functions --> adding or extracting information from and to existing plots
    --> use so called devices (f.e. mouse)

# Univariate continous variable plots

## Base R: How can you create a histogram in R? ✅

using the `hist`function

```{r}
x <- rnorm(100)
hist(x)
```

## Base R: How can you create a boxplot in R? ✅

using the `boxplot` function

```{r}
x <- rnorm(100)
boxplot(x)
```

## Base R: How can you create a density plot? ✅

```{r}
x <- density(rnorm(100))
plot(x)
```

Why does this work?

`density` gives us a density estimate for the variable `x`.
This density estimate has a special class, which has a specific plotting function.
This function, plots the denisty nicely.

```{r}
class(x)
```

## Base R: How do you create a scatter plot (against the index of the vector of the univariate variable)? ✅

```{r}
x <- rnorm(100)
plot(x)
```

## Base R: How do you create a 1D scatter plot? ❌

```{r}
x <- rnorm(50)
stripchart(x)
```

## Base R: How can you create a QQ-plot in R? ✅

```{r}
y <- rgamma(1000, shape = 10)
qqnorm(y)
```

# Univariate categorical variable plot

## Base R: How do you create a histogram for the occurrence of a categorical variable? ✅

```{r}
fac1 <- factor(
    sample(1:3, 15, replace = TRUE),
    labels = c("Pidgey", "Metapod", "Ratata")
)
plot(fac1)
```

## Base R: How do you create a pie-chart of the occurrence for a categorical variable? ❌

```{r}
fac2 <- factor(
    sample(1:3, 15, replace = TRUE),
    labels = c("Dodrio", "Nidoran", "Pidgeot")
)
pie(table(fac2))
```

## What does table do? ❌

takes in categorical variables,
produces frequence/contingency tables.

```{r}
table(fac1)
```

--> frequency of the occurences

```{r}
table(fac1, fac2)
```

--> contingency/cooccurrence table
--> variables must have same length

## Base R: How do you plot the covariation of two variables? ✅

```{r}
x <- rnorm(100)
y <- x * 1.7 + 2 + rnorm(100)
plot(x, y)
```

## Base R: How do you plot the pair-wise scatterplots of a data frame? ❌

for dataframes:

```{r}
x <- rnorm(100)
y <- x * 1.7 + 2 + rnorm(100)
z <- x * y + rnorm(100)
r <- runif(100)
X <- data.frame(x = x, y = y, z = z, r = r)
pairs(X)
```

Can't make this work for matrices tho 🤪

## Base R: Plotting the covariation of two discrete/categorical variables? ❌

```{r}
fac1 <- factor(
    sample(1:2, 15, replace = TRUE),
    labels = c("is alien", "is human")
)
fac2 <- factor(
    sample(1:2, 15, replace = TRUE),
    labels = c("likes pickles", "dislikes pickles")
)
plot(fac1, fac2)
```

## Base R: Plotting a categorical variable against a continuous variable? ❌

```{r}
race <- factor(
    c(rep("is human", 50), rep("is alien", 50))
)
height <- c(rnorm(50, 175, 10), rnorm(50, 140, 30))
plot(height ~ race)
```

Important: The reponse is always first.
Imagine it as: "Model height by race."
So "Model response by variable."
"y ~ x"
👽

If you switch that, you get a "spine-plot" 💀

```{r}
plot(race ~ height)
```

--> category-split per variable expression

## Base R: What are 3 methods for displaying 3-dimensional data ❌ ✅

* `image` --> display as map of pixels, i.e. an image
* `contour` --> display as an image, with "height" line indicators
* `persp` --> display as 3D polygonal shape

## Base: Demonstrate image ❌

```{r}
require(grDevices)
x <- y <- seq(-4 * pi, 4 * pi, len = 200)
r <- sqrt(outer(x^2, y^2, "+"))
z <- cos(r^1) * exp(-r / 6)
image(x, y, z)
```

## Base R: Demonstrate contour ✅

```{r}
contour(x, y, z)
```

## Base R: Demonstrate persp ✅

```{r}
persp(x, y, z)
```

## Base R: What are really cool high level plotting commands that change how these functions work? ❌

* `add = TRUE`: plot commands works like low-level function, i.e. becomes additive
* `axes = FALSE`: don't include axis, good for custom axes
* `log = "x"`, `log = "y"` or `log = "xy"`: log transform a variable
* `xlab = "text"`: change label of x-axis
* `ylab = "text"`: change label of y-axis
* `main = "text"`: add title
* `sub = "text"`: add subtitle
* `type =` type of plot, different options, defaults to `p`
    * `l` lines
    * `b` points connected by lines
    * `o` points overlayed by lines
    * `h` vertical lines to zero axis
    * `s` step function ._| line "follows" along defining point
    
    * `S` step function !‾ line "trails" before defining point
    * `n`: dont show any indicators, helpful for lowe level function

```{r}
plot(x, y, axes = FALSE, xlab="", ylab="", type="b")
```

## Base R: What are useful functions for plotting points, lines and shapes? ❌

* `lines`: plot a line from data points
* `points`: plot points from data points
* `abline`: helper function to plot lines
* `polygon`: add a polygon to an existing plot

## Base R: Demonstrate lines, points and polygon ❌

```{r}
x <- rnorm(20) + 1:20
plot(x, type = "n")
polygon(c(1:20, 20), c(x[], x[1]), col = "#72c78c")
points(x)
lines(x)
```

## Demonstrate different parameters for abline (2 in total) ❌

```{r}
plot(1, 1, type="n", ylim=c(-2, 2), xlim=c(-2, 2)) + abline(v=0) + abline(h=0)
```

## Base R: What are useful functions used for styling plots? ❌

* `title`: add titles (`main`) or subtitles (`sub`)
* `text`: add text add x, y coordinates into the plot
* `axis`: used to specify anything about plot axis

## What are useful parameters for styling plots? ❌
* `fill =`: fill color of shape
* `lty =`: linetype/style of line
* `lwd =`: width of lines
* `col =`: color of line
* `pch =`: style of point --> character i.e. pointcharcter ~ pch

```{r}
plot(1, 1, xlim = c(1, 10), ylim = c(0, 5), type = "n", axes = FALSE)
points(1:9, rep(4.5, 9), cex = 1:9, col = 1:9, pch = 0:8)
text(1:9, rep(3.5, 9), labels = paste(0:9), cex = 1:9, col = 1:9)
points(1:9, rep(2, 9), pch = 9:17)
text((1:9) + 0.25, rep(2, 9), paste(9:17))
points(1:8, rep(1, 8), pch = 18:25)
text((1:8) + 0.25, rep(1, 8), paste(18:25))
```

## Base R: What are two different types of interactive functions in R? ❌

* `locator` --> adding points to graph
* `identify` --> highlight data in a graph

## How do you use locator? ❌

```{r}
x <- rnorm(100)
y <- rnorm(100)
plot(x, y)
while (TRUE) {
    point <- locator(1)
    print(point)
    text(point, "clicked here")
}
```

## How do you use identify? ❌

```{r}
x <- rnorm(100)
y <- rnorm(100)
plot(x, y)
id <- identify(x, y, labels = 1:100)
print(id)
```

## What are graphics devices in R? ✅

They are essentially soft- or hardware entities, that can be the destination for graphical displays.
Devices include:
* Screen devices --> logical entities provided by software like Rstudio, X11 and so on
* File devices --> a file to write to
* Printer device --> print to physical printer, or a pdf
* Other devices --> include html widgetss

## How do you "permanently" change the graphical parameters of a device? ✅

using `par`

```{r}
par(pch = "*")
x <- rnorm(100)
y <- rnorm(100)
plot(x, y)
```

When closing windows between calls, then pars get reset!

```{r}
x <- rnorm(100)
y <- rnorm(100)
plot(x, y)
```

## How do you list all graphical parameters? ✅

using `par()` without parameters

```{r}
par()
```

## What are the most important graphical parameters (according to the lecturers)? ✅

* `pch`: plotting character, either as character string, or as index [0-18]
* `lty`: linetype, supplied as integer [1, ], not specified how many different
* `lwd`: line width as multiples of the default width
* `col`: specifies the color of a graphical element
* `cex`: scaling of displayed characters

## What are parameters concerning the figure margins? ❌

* `mai=(1, 2, 3, 4)`: for margin in inches
* `mar=(1, 2, 3, 4)`: for margin in lines

## What is the association between the margin parameters and their position around the plot? ❌

`mar=(bottom, left, top, right)`

doesnt really make sense

## How can you adjust the outer margin? What's the default? ❌

* `oma=(bottom, left, top, right)`: margin in lines
* `omi=(bottom, left, top, right)`: margin in inches

--> default is 0, distance between blot and its borders

## What are the two functions that enable multiple plots per figure in base R? ❌

* `par` using `mfrow` or `mfcol`
* `layout`


## What is a great way of plotting function curves in R? ❌

```{r}
curve(sin, -2 * pi, 2 * pi)
```

## How do you use par to create multi columns? How do the two ways differ? ❌

mfrow fills row-wise:
1 2
3 4

```{r}
par(mfrow = c(2, 2))
curve(sin, -2 * pi, 2 * pi)
curve(cos, -2 * pi, 2 * pi)
curve(tan, -2 * pi, 2 * pi)
curve(acos, -2 * pi, 2 * pi)
```

mfcol fills column-wise:
1 3
2 4

```{r}
par(mfcol = c(2, 2))
curve(sin, -2 * pi, 2 * pi)
curve(cos, -2 * pi, 2 * pi)
curve(tan, -2 * pi, 2 * pi)
curve(acos, -2 * pi, 2 * pi)
```

## How do you specify complex layouts using the layout function? ❌

creates a grid base dlayout

`layout` has parameters:
(
    `mat`: a matrix, its dimensions specify the layout of the multiplot,
        numbering denotes order of cells
    `widths`: widths of the columns, default evenely spaced
    `heights`: heights of the rows, default evenely spaced
    `respect`: default `FALSE` means stretching and squishing plot to fit cell, 
        if `TRUE` then retain aspect ratio of plot
)

```{r}
layout(
    mat = rbind(c(3, 0), c(1, 2)),
    heights = c(1, 2),
    widths = c(2, 1)
)
curve(tan, -2 * pi, 2 * pi)
curve(sin, -2 * pi, 2 * pi)
curve(cos, -2 * pi, 2 * pi)
```

## How do you list all possibles devices in R? ❌

```{r}
help(Devices)
```

## How can you create a jpeg device? ❌

```{r}
jpeg(
    file = "/home/paul/Downloads/sinus.jpeg",
    width = 150,
    height = 150,
    bg = "white"
)
curve(sin, -2 * pi, 2 * pi)
dev.off()
```

## Why should you use dev.off() after you are done plotting to a device?

* to free up the system resources
* to avoid unexpected behaviour

## What is the basic idea of the grammar of graphics?

* move beyond simple categories of plots lile "scatterplot",
    "lineplot", "pie-chart", etc.
* instead identify structures and conventions that underpin these categories
* learn to recombine them and use them on different data
--> learn base rules and apply them differently like one does with words in a language

## Describe the layered grammer of graphics and give examples for each layer?

* `data`: structured data source
    example: dataframe
* `mapping`/`aesthetic`: relation between data and graphical property
    example: variable "petal_width" maps onto x-coordinate
* `statistical transformation` (optional): most often just the identity, some extra transformation
    example: binning, aggregation, smoothing, kde
* `geom`: representation of variable as some sort of geometric object
    require certain outputs from aesthetic
    example: bars, lines, points
* `scale`: mapping between the data values and their visual representations
    its the transformation from one dimension into the other: brain size --> color
    example: map numeric variable to color gradient
* `coordinates`: which coordinate system to use
    example: cartesian, polar
* `faceting`: splitting the data, subsettign it, different display based on factes
    example: hue in ggplot2


## Why is data its own layer?

* because a visualization can be independent of the data it represents.
* a given visualization can be applied to many different data sets.

## What are some aesthetics to map to (about 5 examples)?

* x and y axis
* color and alpha
* line style, width
* shape
* size


## How do you plot windows?

```{r}
library(ggplot2)
library(tidyverse)

win <- tibble(
    name = c("1.0", "2.0", "3.1", "95", "98", "2000", "XP", "7", "8", "10"),
    version = as.integer(name)
)

ggplot(win, aes(x = name, y = version, hue = name)) +
    geom_col()
```

## How are statistics and geoms related

* statistics and geoms can be linked
example:
--> group and count statistic is linked to bar charts and vice versa

## How do you load the penguins data set?

```{r}
library("palmerpenguins")
```

## Plot a histogram of the body weight of the penguins

```{r}
ggplot(penguins, aes(x = body_mass_g)) +
    geom_histogram()
```

## Plot a histogram of the body weight of the penguins with a bin width of 500

```{r}
ggplot(penguins, aes(x = body_mass_g)) +
    geom_histogram(binwidth = 500)
```

## Now split and stack the histogram by penguin species

```{r}
ggplot(penguins, aes(x = body_mass_g, fill = species)) +
    geom_histogram(binwidth = 500)
```

## Noe create the histogram without stacking the different species values, but displaying them next to eahc other

```{r}
ggplot(penguins, aes(x = body_mass_g, fill = species)) +
    geom_histogram(binwidth = 500, position = "dodge")
```

## Use faceting to create a colored histogram of the bodyweight for each species

```{r}
ggplot(penguins, aes(x = body_mass_g, fill = species)) +
    geom_histogram(binwidth = 500) +
    facet_wrap(~species)
```

## Now show the proportion of penguin species by island as a stacked bar plot

```{r}
ggplot(penguins, aes(x = island, fill = species)) +
    geom_bar()
```

## Now show the distibution of body weights per species in boxplots

```{r}
ggplot(penguins, aes(x = species, y = body_mass_g)) +
    geom_boxplot()
```

## Now show this exact box plot, but split and color groups for each species also by gender

```{r}
ggplot(penguins, aes(x = species, y = body_mass_g, color = sex)) +
    geom_boxplot()
```

## show a scatter plot contrasting body mass and flipper length of the penguins

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
    geom_point()
```

## Now make this same scatterplot and distinguish the penguins species by color and sex by symbol

```{r}
ggplot(penguins, aes(
    x = flipper_length_mm,
    y = body_mass_g,
    color = species,
    shape = sex
)) +
    geom_point()
```

## How do you show both the base histogram of the bodyweights but also overlay the species specific histograms

```{r}
ggplot(penguins, aes(
    x = body_mass_g,
)) +
    geom_histogram() +
    geom_histogram(aes(color = species),
        fill = NA,
        position = "identity"
    )
```

## What is the effect of using position=identity in a hsitogram?

bars arent being stacked over each other when split by some variable occurs,
rather than that the bars are displayed as overlapping elements

## Draw a plot describing the distribution of body weights, both as a histogram and as a density plot

```{r}
ggplot(penguins, aes(x = body_mass_g, y = stat(density))) +
    geom_histogram(fill = "grey") +
    geom_density()
```

## How can you set the title and axis labels of a plot, that has been facet wrapped

```{r}
ggplot(penguins, aes(x = body_mass_g, fill = species)) +
    geom_histogram(binwidth = 500) +
    facet_wrap(~species) +
    ggtitle("Body mass distribution by species") +
    xlab("Body mass gramm") +
    ylab("Counts")
```

## How do you set the title of a plot in ggplot

using the low-level plotting funtion `ggtitle`

## Given the facet wrapped histogram, how could you draw the mean body weight across the entire data set onto each facet?

need to create an extra data frame --> mean is the same per species

```{r}
pengu_mean <- tibble(
    species = unique(penguins$species),
    mean_bwg = rep(mean(penguins$body_mass_g, na.rm = TRUE), 3)
)

ggplot(penguins, aes(x = body_mass_g, fill = species)) +
    geom_histogram(binwidth = 500) +
    geom_vline(data = pengu_mean, aes(xintercept = mean_bwg)) +
    facet_wrap(~species)
```

## Okay this one is going to be complicated
## Lets say you want a histogram facet wrapped by species
## The catch is that overlayed over each histogram you want the density of the entire data set
## consider again the body mass variable

need a really weird data frame here.
remember, each facet needs to receive the same density.

```{r}
pengu_dens <- tibble(
    species = rep(unique(penguins$species), each = nrow(penguins)),
    body_mass_g = rep(penguins$body_mass_g, 3)
)

ggplot(penguins, aes(x = body_mass_g, y = stat(density))) +
    geom_histogram(binwidth = 500, aes(fill = species)) +
    geom_density(data = pengu_dens) +
    facet_wrap(~species)
```

## In the gganimate package what can you use the "transition_state" function for?

you can set up multiple states for a plot and switch between them

```{r}
library("gganimate")

ggplot(penguins, aes(x = body_mass_g, y = stat(density))) +
    geom_histogram(binwidth = 500, aes(fill = species)) +
    geom_density(data = pengu_dens) +
    transition_states(species)
```

## How can you update the user on the current frame of the transition state?

Using variable injection in a text attribute:
`{closest_state}`
`{frame}`
`{nframes}`

```{r}
library("gganimate")

ggplot(penguins, aes(x = body_mass_g, y = stat(density))) +
    geom_histogram(binwidth = 500, aes(fill = species)) +
    geom_density(data = pengu_dens) +
    transition_states(species) +
    ggtitle("Now showing {closest_state}",
        subtitle = "Frame {frame} of {nframes}"
    )
```

## How can you fine tune the animation that is used for data fading?

* `enter_` functions like `enter_fade`
* `exit_` functions like `exit_shrink`

## Give an example for using the gapminder data how to animate time

```{r}
library("gapminder")

ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, size = pop, colour = country)) +
    geom_point(show.legend = FALSE, alpha = 0.7) +
    scale_color_viridis_d() +
    scale_size(range = c(2, 12)) +
    scale_x_log10() +
    labs(x = "GDP per capita", y = "Life expectancy") +
    transition_time(year) +
    labs(title = "Year: {frame_time}")
```

## How do create an interactive graph using plotly?

```{r}
library(plotly)

p <- ggplot(penguins, aes(flipper_length_mm, body_mass_g, color = species)) +
    geom_point()

pltly <- ggplotly(p, width = 900, height = 900)
pltly
```

## How can you customize plotly plots?

using:
* `layout()`
* `style()`

with the dplyr notation

## What are possible attributes for/ capabilities of layout?

* `hoverlabel`: to stylize hoverlabels
* `dragmode`: to change drag benhavior
* `selectdirection`: to constrain rectangular selections
* add dropdowns

## What are some of the capabilties of style?

* control tooltip content
* turn hovering on/off 
* add marker points to lines